The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2024-02-28 r_r/submission/manuscript_rihe_r_r.tex
knit pdf

%!TEX TS-program = xelatex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\title{The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities}
\author{}
\date{\vspace{-2.5em}}

@@

  \usepackage{unicode-math}

@@

@@

\urlstyle{same} % disable monospaced font for URLs

@@

@@

We situate this study within scholarship on nonresident enrollment at public universities. Author (XXXXa) analyzed recruiting visits to public high schools made by 15 public research universities in 2017. The authors found that 12 of the 15 universities made more visits to out-of-state schools than in-state schools. These out-of-state visits focused on affluent, predominantly white public schools. Set against the scholarship about declining state appropriations and growing nonresident enrollment (Author, XXXXb, XXXXc), these results suggest that many public research universities prioritize enrollment from wealthy out-of-state students who can afford to pay nonresident tuition prices.

A separate, multidisciplinary literature has analyzed linkages between private school and college access (Bryk, Lee, \& Holland, 1993; Chetty, Deming, \& Friedman, 2023; Khan, 2011; Kingston \& Lewis, 1990). This research observes strong links -- and student flows -- between private high schools and selective private universities. Private schools enroll many students who can afford private university tuition (Stevens, 2007). To date, scholarship on nonresident enrollment by public universities has progressed in isolation from scholarship on links between private schools and colleges.

@@

Author (XXXXb) draw from resource dependence theory to conceptualize the relationship between state appropriations and nonresident enrollment. Organizations are dependent on external resources that are important for organizational survival/stability and cannot be easily replaced. Public universities depend on state appropriations but these funds became increasingly uncertain in the 1980s and 1990s (Delaney \& Doyle, 2011). Resource dependence theory states that one response to resource decline or uncertainty is to seek alternative, substitute resources (Pfeffer \& Salancik, 1978). Tuition revenue is a potential substitute for state appropriations. Most states cap resident tuition prices, but not nonresident tuition prices. Therefore, Author (XXXXb) argued that public universities have a financial incentive to grow nonresident enrollment in response to declines in state appropriations. Analyzing panel data from 2002-03 to 2012-13, Author (XXXXb) found that a 10\% decline in state appropriations was associated with a 2.7\% increase in nonresident enrollment at all public universities and a 5.0\% increase in nonresident enrollment at research-extensive universities.

Other studies of nonresident enrollment examine university enrollment composition. Using data from NPSAS 2011-12, Author (XXXXc) show that compared to resident students at public research universities, nonresident students generate higher net tuition revenue, are more affluent, are less likely to be Pell recipients, and are less likely to be Black or LatinX. Institution-level panel models reveal that growth in the share of nonresident students was associated with a decline in the share of Pell recipients and a decline in the share of Black and LatinX students. Author (XXXXd) examine whether growth in the number of nonresident students causes resident enrollment to decline. They find that nonresident enrollment ``crowds out'' resident enrollment at prestigious public research universities, but not at less prestigious public research universities.

@@

Earlier in the ``enrollment funnel,'' universities utilize marketing and recruiting interventions to attract nonresident applicants. One widely utilized intervention is off-campus recruiting visits by admissions counselors to local high schools. Author (XXXXa) analyzed off-campus recruiting visits to public high schools by 15 public research universities. Recruiting visits to in-state high schools tended to exhibit modest racial and socioeconomic bias. However, 12 of the 15 universities made more visits to out-of-state high schools than in-state high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Salazar (2022) conducted geospatial analyses of out-of-state recruiting visits to Los Angeles and Dallas by four public research universities. She found that universities engage in ``recruitment redlining -- the circuitous avoidance of predominantly Black and Latinx communities along recruiting visit paths'' (p.~585). Although Author (XXXXa) and Salazar (2022) focused on visits to public high schools, prior research has not considered visits to private schools by public research universities. For example, what is the scale of these visits and what are the characteristics of visited schools?

@@

Although the literature on nonresident enrollment developed in isolation from the literature on private school to private university linkages, both phenomena are part of a broader story about the changing nature of school segregation and the competition for affluent, traditional college students. Drawing from resource dependence theory, as public research universities become more reliant, they have a financial incentive to target out-of-state private schools, which tend to enroll affluent students that can pay high tuition prices. Although a handful of public research universities have relied on nonresident enrollment for many decades (e.g., University of Michigan, University of Delaware), public research universities began substantially increasing nonresident enrollment en masse around 2005 (Author, XXXXb). By contrast, selective private universities have long relied on enrollment from private schools (Persell \& Cookson, 1985; Stevens, 2007). Therefore, the pursuit of out-of-state private school students by public research universities is a relatively recent encroachment on the historic territory of selective private universities and in-state flagships. Speaking about this competition, an enrollment management consultant said, ``\,`Everybody wants to go to the magic island of full pay students, but it's rapidly shrinking real estate'\,'' (Rivard, 2013). By analyzing the network of recruiting visits, we show whether public research universities and selective private universities are competing for students at similar sets of private schools. This may be direct competition for the same student or vying to be the second choice for students who are rejected by their top choice.

@@

For public universities in our analysis sample, we assessed the quality of the web-scraped data by asking the enrollment/admissions office for these data. If a university denied or ignored our request, we issued a formal public records request. Of the 15 public universities analyzed in this study, we received requested data from 11 universities. For these 11 universities, web-scraped data was very similar to publicly requested data. For universities that sent us data, the analyses below use ``requested'' rather than ``scraped'' data because requested data tended to have more events. Broad patterns were similar across requested data versus scraped data, and results based on scraped data are available upon request. Author (XXXXa) describe data quality checks in greater detail.

@@

This study engages with two, previously disconnected literatures. Scholarship on nonresident enrollment observes that many public research universities began growing nonresident enrollment following declines in state appropriations (Author, XXXXb). Given that nonresident tuition price is often two- to three-times resident price, it is unsurprising that nonresident students tend to be more affluent than resident students, are less likely to be first-generation students, and are less likely to be Black or Latinx (Author, XXXXc). Public research universities recruit nonresident students using institutional financial aid packages (Leeds \& DesJardins, 2015) and through recruiting interventions that target earlier stages in the enrollment funnel. Using nearly the same sample of 15 public research universities used in this study, Author (XXXXa) found that 12 universities made more out-of-state visits than in-state visits and visits to out-of-state public high schools focused on affluent, predominantly white schools. Public research universities have a financial incentive to target private school students, who tend to come from affluent households that are able to pay nonresident tuition price without need-based aid. This study extends the literature on nonresident students by analyzing the scale of visits to private schools, the overlap and divergence in recruiting networks between public and private universities, and the characteristics of visited private schools.

@@

Considering the practical importance of off-campus recruiting visits, what are the practical implications of our research? Prior scholarly and market research shows that students are sensitive to recruiting interventions, including recruiting visits (Holland, 2019; Ruffalo Noel-Levitz, 2018). Consequently, the racial and socioeconomic composition of the schools that a university chooses to visit affects the racial and socioeconomic composition of the university. Our research on visits to public high schools (Author, XXXXa) and this study on visits to private schools find that our sample of public research universities and selective private universities tend to focus their recruiting efforts on affluent and predominantly white schools and communities. Some universities may publicly proclaim a commitment to social mobility and privately pursue an intentional recruiting strategy that targets affluent schools and communities. We offer practical guidance for universities that may be unintentionally implementing a recruiting strategy that privileges the privileged.

@@

We suspect that many universities may have incomplete knowledge about which schools they recruit from and how these schools compare to those they do not recruit from. Our rationale for this statement is that employee turnover in the enrollment management profession is quite high (Hoover, 2023) and many universities rely on expertise from third-party enrollment management consultancies (Author, XXXXe). Consistent with this idea, when issued public records requests to collect data on ``student list'' purchases from about 100 public universities (Author, XXXXf), we found that many public universities did not know who or how they purchased student lists, typically because the person who did this had left the university and/or the job was outsourced to a consultancy.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-RN4758}{}}%
Author. (XXXXa).

\leavevmode\vadjust pre{\hypertarget{ref-list_biz}{}}%
Author. (XXXXe).

\leavevmode\vadjust pre{\hypertarget{ref-list_empirics}{}}%
Author. (XXXXf).

\leavevmode\vadjust pre{\hypertarget{ref-RN3753}{}}%
Author. (XXXXb).

\leavevmode\vadjust pre{\hypertarget{ref-RN3685}{}}%
Author. (XXXXc).

\leavevmode\vadjust pre{\hypertarget{ref-RN4290}{}}%
Author. (XXXXd).

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN1791}{}}%
Curs, B. R., \& Singell, L. D. (2002). An analysis of the application and enrollment processes for in-state and out-of-state students at a large public university. \emph{Economics of Education Review}, \emph{21}(2), 111--124. Retrieved from \url{\%3CGo\%20to\%20ISI\%3E://000173740100002}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN4625}{}}%
Hill, L. D. (2008). School strategies and the "college-linking" process: Reconsidering the effects of high schools on college enrollment. \emph{Sociology of Education}, \emph{81}(1), 53--76. Journal Article. Retrieved from \url{\%3CGo\%20to\%20ISI\%3E://WOS:000256634400003}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN4495}{}}%
Jack, A. A. (2019). \emph{The privileged poor: How elite colleges are failing disadvantaged students} (p. 276). Cambridge, Massachusetts: Harvard University Press,. Retrieved from \url{Restricted\%20to\%20UCLA.\%20Limited\%20to\%20three\%20concurrent\%20users.\%20Try\%20again\%20later\%20if\%20unavailable.\%20http://openurl.cdlib.org?sid=UCLA:CAT\&genre=book\&__char_set=utf8\&isbn=9780674239647}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN3522}{}}%
Killgore, L. (2009). Merit and competition in selective college admissions. \emph{Review of Higher Education}, \emph{32}(4), 469--488. Retrieved from \url{\%3CGo\%20to\%20ISI\%3E://WOS:000266737500002}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN1722}{}}%
McPherson, M., Smith-Lovin, L., \& Cook, J. M. (2001). Birds of a feather: Homophily in social networks. \emph{Annual Review of Sociology}, \emph{27}, 415--444. Journal Article. Retrieved from \url{\%3CGo\%20to\%20ISI\%3E://000170748100017}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN1549}{}}%
Winston, G. C. (1999). Subsidies, hierarchy and peers: The awkward economics of higher education. \emph{Journal of Economic Perspectives}, \emph{13}(1), 13--36. Retrieved from \url{\%3CGo\%20to\%20ISI\%3E://000078904300002}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN4758}{}}%
Author. (XXXXa).

\leavevmode\vadjust pre{\hypertarget{ref-list_biz}{}}%
Author. (XXXXe).

\leavevmode\vadjust pre{\hypertarget{ref-list_empirics}{}}%
Author. (XXXXf).

\leavevmode\vadjust pre{\hypertarget{ref-RN3753}{}}%
Author. (XXXXb).

\leavevmode\vadjust pre{\hypertarget{ref-RN3685}{}}%
Author. (XXXXc).

\leavevmode\vadjust pre{\hypertarget{ref-RN4290}{}}%
Author. (XXXXd).

@@

\begin{table}[!h]

\caption{\label{tab:sample-characteristics}Characteristics of universities in analysis sample}
\centering
\resizebox{\linewidth}{!}{
\fontsize{13}{15}\selectfont
\begin{tabular}[t]{l>{\raggedright\arraybackslash}m{4.5cm}>{\raggedleft\arraybackslash}m{1.8cm}>{\raggedleft\arraybackslash}m{3.8cm}>{\raggedleft\arraybackslash}m{3.8cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.7cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{3cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{3cm}}
\toprule
Classification & University & Rank & 25th Percentile SAT/ACT Composite Score & 75th Percentile SAT/ACT Composite Score & In-State Tuition + Fees & Out-of-State Tuition + Fees & Total Enrolled Freshmen & Percent Out-of-State Freshmen & Percent Pell Recipients & Percent White Freshmen & Percent Black Freshmen & Percent Asian Freshmen & Percent Hispanic Freshmen & Percent International Freshmen\\
\midrule
 & UC Berkeley & 22 & 1,316 & 1,527 & \$13,807 & \$41,076 & 6,252 & 24.4\% & 19.4\% & 25.6\% & 1.8\% & 42.7\% & 13.8\% & 9.3\%\\

@@

\begin{table}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}

@@

  \usepackage{unicode-math} % this also loads fontspec

@@

\ifPDFTeX\else
  % xetex/luatex font selection

@@

@@

\urlstyle{same}

\title{The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities}
\author{}
\date{\vspace{-2.5em}}

@@

We situate this study within scholarship on nonresident enrollment at public universities. Salazar, Jaquette, \& Han (2021) analyzed recruiting visits to public high schools made by 15 public research universities in 2017. The authors found that 12 of the 15 universities made more visits to out-of-state schools than in-state schools. These out-of-state visits focused on affluent, predominantly white public schools. Set against the scholarship about declining state appropriations and growing nonresident enrollment (Jaquette \& Curs, 2015; Jaquette, Curs, \& Posselt, 2016), these results suggest that many public research universities prioritize enrollment from wealthy out-of-state students who can afford to pay nonresident tuition prices.

A separate, multidisciplinary literature has analyzed linkages between private school and college access (Bryk, Lee, \& Holland, 1993; Chetty, Deming, \& Friedman, 2023; Khan, 2011; Kingston \& Lewis, 1990; Stevens, 2007). This research observes strong links -- and student flows -- between private high schools and selective private universities. To date, scholarship on nonresident enrollment by public universities has progressed in isolation from scholarship on links between private schools and colleges.

@@

Jaquette \& Curs (2015) draw from resource dependence theory to conceptualize the relationship between state appropriations and nonresident enrollment. Organizations are dependent on external resources that are important for organizational survival/stability and cannot be easily replaced. Public universities depend on state appropriations but these funds became increasingly uncertain in the 1980s and 1990s (Delaney \& Doyle, 2011). Resource dependence theory states that one response to resource decline or uncertainty is to seek alternative, substitute resources (Pfeffer \& Salancik, 1978). Tuition revenue is a potential substitute for state appropriations. Most states cap resident tuition prices, but not nonresident tuition prices. Therefore, Jaquette \& Curs (2015) argued that public universities have a financial incentive to grow nonresident enrollment in response to declines in state appropriations. Analyzing panel data from 2002-03 to 2012-13, Jaquette \& Curs (2015) found that a 10\% decline in state appropriations was associated with a 2.7\% increase in nonresident enrollment at all public universities and a 5.0\% increase in nonresident enrollment at research-extensive universities.

Other studies of nonresident enrollment examine university enrollment composition. Using data from NPSAS 2011-12, Jaquette et al. (2016) show that compared to resident students at public research universities, nonresident students generate higher net tuition revenue, are more affluent, are less likely to be Pell recipients, and are less likely to be Black or LatinX. Institution-level panel models reveal that growth in the share of nonresident students was associated with a decline in the share of Pell recipients and a decline in the share of Black and LatinX students. Curs \& Jaquette (2017) examine whether growth in the number of nonresident students causes resident enrollment to decline. They find that nonresident enrollment ``crowds out'' resident enrollment at prestigious public research universities, but not at less prestigious public research universities.

@@

Earlier in the ``enrollment funnel,'' universities utilize marketing and recruiting interventions to attract nonresident applicants. One widely utilized intervention is off-campus recruiting visits by admissions counselors to local high schools. Salazar et al. (2021) analyzed off-campus recruiting visits to public high schools by 15 public research universities. Recruiting visits to in-state high schools tended to exhibit modest racial and socioeconomic bias. However, 12 of the 15 universities made more visits to out-of-state high schools than in-state high schools. These out-of-state visits focused on public schools in affluent, predominantly white communities. Salazar (2022) conducted geospatial analyses of out-of-state recruiting visits to Los Angeles and Dallas by four public research universities. She found that universities engage in ``recruitment redlining -- the circuitous avoidance of predominantly Black and Latinx communities along recruiting visit paths'' (p.~585). Although Salazar et al. (2021) and Salazar (2022) focused on visits to public high schools, prior research has not considered visits to private schools by public research universities. For example, what is the scale of these visits and what are the characteristics of visited schools?

@@

Although the literature on nonresident enrollment developed in isolation from the literature on private school to private university linkages, both phenomena are part of a broader story about the changing nature of school segregation and the competition for affluent, traditional college students. Drawing from resource dependence theory, as public research universities become more tuition reliant, they have a financial incentive to target out-of-state private schools, which tend to enroll affluent students that can pay high tuition prices. Although a handful of public research universities have relied on nonresident enrollment for many decades (e.g., University of Michigan, University of Delaware), public research universities began substantially increasing nonresident enrollment en masse around 2005 (Jaquette \& Curs, 2015). By contrast, selective private universities have long relied on enrollment from private schools (Persell \& Cookson, 1985; Stevens, 2007). Therefore, the pursuit of out-of-state private school students by public research universities is a relatively recent encroachment on the historic territory of selective private universities and in-state flagships. Speaking about this competition, an enrollment management consultant said, ``\,`Everybody wants to go to the magic island of full pay students, but it's rapidly shrinking real estate'\,'' (Rivard, 2013). By analyzing the network of recruiting visits, we show whether public research universities and selective private universities are competing for students at similar sets of private schools. This may be direct competition for the same student or vying to be the second choice for students who are rejected by their top choice.

@@

For public universities in our analysis sample, we assessed the quality of the web-scraped data by asking the enrollment/admissions office for these data. If a university denied or ignored our request, we issued a formal public records request. Of the 15 public universities analyzed in this study, we received requested data from 11 universities. For these 11 universities, web-scraped data was very similar to publicly requested data. For universities that sent us data, the analyses below use ``requested'' rather than ``scraped'' data because requested data tended to have more events. Broad patterns were similar across requested data versus scraped data, and results based on scraped data are available upon request. Salazar et al. (2021) describe data quality checks in greater detail.

@@

This study engages with two, previously disconnected literatures. Scholarship on nonresident enrollment observes that many public research universities began growing nonresident enrollment following declines in state appropriations (Jaquette \& Curs, 2015). Given that nonresident tuition price is often two- to three-times resident price, it is unsurprising that nonresident students tend to be more affluent than resident students, are less likely to be first-generation students, and are less likely to be Black or Latinx (Jaquette et al., 2016). Public research universities recruit nonresident students using institutional financial aid packages (Leeds \& DesJardins, 2015) and through recruiting interventions that target earlier stages in the enrollment funnel. Using nearly the same sample of 15 public research universities used in this study, Salazar et al. (2021) found that 12 universities made more out-of-state visits than in-state visits and visits to out-of-state public high schools focused on affluent, predominantly white schools. Public research universities have a financial incentive to target private school students, who tend to come from affluent households that are able to pay nonresident tuition price without need-based aid. This study extends the literature on nonresident students by analyzing the scale of visits to private schools, the overlap and divergence in recruiting networks between public and private universities, and the characteristics of visited private schools.

@@

Considering the practical importance of off-campus recruiting visits, what are the practical implications of our research? Prior scholarly and market research shows that students are sensitive to recruiting interventions, including recruiting visits (Holland, 2019; Ruffalo Noel-Levitz, 2018). Consequently, the racial and socioeconomic composition of the schools that a university chooses to visit affects the racial and socioeconomic composition of the university. Our research on visits to public high schools (Salazar et al., 2021) and this study on visits to private schools find that our sample of public research universities and selective private universities tend to focus their recruiting efforts on affluent and predominantly white schools and communities. Some universities may publicly proclaim a commitment to social mobility and privately pursue an intentional recruiting strategy that targets affluent schools and communities. We offer practical guidance for universities that may be unintentionally implementing a recruiting strategy that privileges the privileged.

@@

We suspect that many universities may have incomplete knowledge about which schools they recruit from and how these schools compare to those they do not recruit from. Our rationale for this statement is that employee turnover in the enrollment management profession is quite high (Hoover, 2023) and many universities rely on expertise from third-party enrollment management consultancies (Jaquette, Salazar, \& Martin, 2022). Consistent with this idea, when issued public records requests to collect data on ``student list'' purchases from about 100 public universities (Salazar, Jaquette, \& Han, 2022), we found that many public universities did not know who or how they purchased student lists, typically because the person who did this had left the university and/or the job was outsourced to a consultancy.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-RN4953}{}}%
Borgatti, S. P. (1998). Ego networks. Retrieved from \url{http://www.analytictech.com/networks/egonet.htm}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN4290}{}}%
Curs, B. R., \& Jaquette, O. (2017). Crowded out? The effect of nonresident enrollment on resident access to public research universities. \emph{Educational Evaluation and Policy Analysis}, \emph{39}(4), 644--669. Journal Article. \url{https://doi.org/10.3102/0162373717704719}

\leavevmode\vadjust pre{\hypertarget{ref-RN1791}{}}%
Curs, B. R., \& Singell, L. D. (2002). An analysis of the application and enrollment processes for in-state and out-of-state students at a large public university. \emph{Economics of Education Review}, \emph{21}(2), 111--124. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://000173740100002}{\textless Go to ISI\textgreater://000173740100002}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN4625}{}}%
Hill, L. D. (2008). School strategies and the "college-linking" process: Reconsidering the effects of high schools on college enrollment. \emph{Sociology of Education}, \emph{81}(1), 53--76. Journal Article. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://WOS:000256634400003}{\textless Go to ISI\textgreater://WOS:000256634400003}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN4495}{}}%
Jack, A. A. (2019). \emph{The privileged poor: How elite colleges are failing disadvantaged students} (p. 276). Cambridge, Massachusetts: Harvard University Press,. Retrieved from \href{Restricted\%20to\%20UCLA.\%20Limited\%20to\%20three\%20concurrent\%20users.\%20Try\%20again\%20later\%20if\%20unavailable.\%20http://openurl.cdlib.org?sid=UCLA:CAT\&genre=book\&__char_set=utf8\&isbn=9780674239647}{Restricted to UCLA. Limited to three concurrent users. Try again later if unavailable. http://openurl.cdlib.org?sid=UCLA:CAT\&genre=book\&\_\_char\_set=utf8\&isbn=9780674239647}

\leavevmode\vadjust pre{\hypertarget{ref-RN3753}{}}%
Jaquette, O., \& Curs, B. R. (2015). Creating the out-of-state university: Do public universities increase nonresident freshman enrollment in response to declining state appropriations? \emph{Research in Higher Education}, \emph{56}(6), 535--565. \url{https://doi.org/10.1007/s11162-015-9362-2}

\leavevmode\vadjust pre{\hypertarget{ref-RN3685}{}}%
Jaquette, O., Curs, B. R., \& Posselt, J. R. (2016). Tuition rich, mission poor: Nonresident enrollment growth and the. \emph{Journal of Higher Education}, \emph{87}(5), 635--673. \url{https://doi.org/10.1353/jhe.2016.0025}

\leavevmode\vadjust pre{\hypertarget{ref-list_biz}{}}%
Jaquette, O., Salazar, K. G., \& Martin, P. (2022). \emph{The student list business: Primer and market dynamics}. The Institute for College Access; Success. Retrieved from \url{https://ticas.org/wp-content/uploads/2022/09/The-Student-List-Business_-Primer-and-Market-Dynamics.pdf}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN3522}{}}%
Killgore, L. (2009). Merit and competition in selective college admissions. \emph{Review of Higher Education}, \emph{32}(4), 469--488. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://WOS:000266737500002}{\textless Go to ISI\textgreater://WOS:000266737500002}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN1722}{}}%
McPherson, M., Smith-Lovin, L., \& Cook, J. M. (2001). Birds of a feather: Homophily in social networks. \emph{Annual Review of Sociology}, \emph{27}, 415--444. Journal Article. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://000170748100017}{\textless Go to ISI\textgreater://000170748100017}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN4758}{}}%
Salazar, K. G., Jaquette, O., \& Han, C. (2021). Coming soon to a neighborhood near you? Off-campus recruiting by public research universities. \emph{American Educational Research Journal}, \emph{58}(6), 1270--1314. \url{https://doi.org/10.3102/00028312211001810}

\leavevmode\vadjust pre{\hypertarget{ref-list_empirics}{}}%
Salazar, K. G., Jaquette, O., \& Han, C. (2022). \emph{Geodemographics of student list purchases: A first look}. The Institute for College Access; Success. Retrieved from \url{https://ticas.org/wp-content/uploads/2022/09/Geodemographics-of-Student-List-Purchases_A-First-Look.pdf}

@@

\leavevmode\vadjust pre{\hypertarget{ref-RN1549}{}}%
Winston, G. C. (1999). Subsidies, hierarchy and peers: The awkward economics of higher education. \emph{Journal of Economic Perspectives}, \emph{13}(1), 13--36. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://000078904300002}{\textless Go to ISI\textgreater://000078904300002}

@@

@@

\begin{table}[!h]
\centering
\caption{\label{tab:sample-characteristics}Characteristics of universities in analysis sample}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{13}{15}\selectfont
\begin{tabular}[t]{l>{\raggedright\arraybackslash}m{4.5cm}>{\raggedleft\arraybackslash}m{1.8cm}>{\raggedleft\arraybackslash}m{3.8cm}>{\raggedleft\arraybackslash}m{3.8cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.7cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{3cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{2.4cm}>{\raggedleft\arraybackslash}m{3cm}}
\toprule
Classification & University & Rank & 25th Percentile SAT/ACT Composite Score & 75th Percentile SAT/ACT Composite Score & In-State Tuition + Fees & Out-of-State Tuition + Fees & Total Enrolled Freshmen & Percent Out-of-State Freshmen & Percent Pell Recipients & Percent White Freshmen & Percent Black Freshmen & Percent Asian Freshmen & Percent Hispanic Freshmen & Percent International Freshmen\\
\midrule
 & UC Berkeley & 22 & 1,316 & 1,527 & \$13,807 & \$41,076 & 6,252 & 24.4\% & 19.4\% & 25.6\% & 1.8\% & 42.7\% & 13.8\% & 9.3\%\\

@@

\begin{table}
\centering
\caption{\label{tab:sample-population-characteristics}Characteristics of universities in analysis sample compared to data collection sample}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{13}{15}\selectfont
\begin{tabular}[t]{lrrrrrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Public research universities} & \multicolumn{6}{c}{Selective private universities} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13}
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Analysis sample (N=15)} & \multicolumn{3}{c}{Data collection sample (N=102)} & \multicolumn{3}{c}{Analysis sample (N=14)} & \multicolumn{3}{c}{Data collection sample (N=57)} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
  & 25th percentile & 50th percentile & 75th percentile & 25th percentile & 50th percentile & 75th percentile & 25th percentile & 50th percentile & 75th percentile & 25th percentile & 50th percentile & 75th percentile\\
\midrule
25th Percentile SAT/ACT Composite Score & 1,060 & 1,126 & 1,164 & 984 & 1,061 & 1,135 & 1,185 & 1,278 & 1,314 & 1,191 & 1,289 & 1,391\\
75th Percentile SAT/ACT Composite Score & 1,292 & 1,332 & 1,366 & 1,204 & 1,282 & 1,350 & 1,373 & 1,433 & 1,496 & 1,375 & 1,472 & 1,542\\
In-State Tuition + Fees & \$10,741 & \$11,785 & \$13,913 & \$9,100 & \$10,750 & \$13,322 & \$47,126 & \$50,140 & \$51,848 & \$47,020 & \$49,914 & \$51,661\\
Out-of-State Tuition + Fees & \$26,866 & \$30,502 & \$34,383 & \$23,433 & \$27,589 & \$32,989 & \$47,126 & \$50,140 & \$51,848 & \$47,020 & \$49,914 & \$51,661\\
Total Enrolled Freshmen & 4,916 & 5,433 & 6,336 & 3,416 & 4,528 & 5,850 & 1,368 & 1,767 & 2,000 & 1,305 & 1,585 & 2,005\\
Percent Out-of-State Freshmen & 21.1\% & 26.9\% & 45.0\% & 15.4\% & 25.8\% & 39.8\% & 63.1\% & 73.6\% & 82.4\% & 66.4\% & 76.6\% & 87.6\%\\
Percent Pell Recipients & 19.5\% & 21.5\% & 26.8\% & 20.3\% & 25.5\% & 33.3\% & 10.2\% & 12.4\% & 16.9\% & 12.8\% & 14.2\% & 16.5\%\\
Percent White Freshmen & 31.6\% & 68.0\% & 75.2\% & 45.3\% & 61.7\% & 72.6\% & 56.5\% & 65.9\% & 71.7\% & 41.1\% & 49.3\% & 63.8\%\\
Percent Black Freshmen & 3.4\% & 4.3\% & 6.6\% & 3.1\% & 5.6\% & 8.9\% & 4.3\% & 4.7\% & 5.0\% & 4.0\% & 5.3\% & 7.1\%\\
Percent Asian Freshmen & 3.4\% & 9.3\% & 29.9\% & 3.1\% & 6.6\% & 13.2\% & 5.5\% & 7.0\% & 14.3\% & 5.9\% & 13.2\% & 18.8\%\\
Percent Hispanic Freshmen & 5.5\% & 8.5\% & 12.3\% & 5.5\% & 8.6\% & 16.8\% & 7.6\% & 10.7\% & 12.8\% & 9.0\% & 10.7\% & 13.3\%\\
Percent International Freshmen & 2.0\% & 4.6\% & 8.7\% & 1.6\% & 3.1\% & 7.1\% & 3.8\% & 5.7\% & 9.0\% & 6.0\% & 9.8\% & 12.3\%\\
\bottomrule
\end{tabular}}
\end{table}
